\subsection{Games and Optimization Problems}

Humans have been playing games for a variety of reasons including
entertainment, sozialization and education for over thousands of
years~\cite{HistoryGames2019}. They are an integral part to both human
culture, society and problem specific understanding. In modern,
recent, times, games and their theory have had an impact on other
fields of research as well. John Nash
proved~\cite{NashEquilibrium2019} that finite games with multiple
players have an optimal strategy, the Nash equilibrium. It is now an
essential part of economics and politics as
well~\cite{NashEquilibrium2019}; most problems with strategic
interaction between several agents can theoretically be solved so that
every agent gets the maximum reward based on every other agent's
optimal decision.

The video game industry in particular has a large impact on the
economy around the world. In 2018, the US video game industry even
matched the US film industry, a much older and more mainstream
industry, in revenue~\cite{VideoGameIndustry2019}. Possibly in part
due to its economic impact, the video game industry has become
responsible for many advances in technology, improving and developing
techniques and hardware~\cite{VideoGameIndustry2019} which have now
become the norm in many other fields as well due to their
effectiveness. A very practical and relevant example is the use of
graphical processing units (GPUs) to speed up the training of machine
learning models.

We have established that games are playing an important part in
society, furthering research and innovation in various areas and
giving people the ability to learn and sozialize in a comfortable
environment. Due to the infinite possibilities in game environments,
an enjoyable game exists for almost everyone~-- similar to how almost
every person finds at least one music genre they enjoy. \\
However, while playing a game to optimize its underlying mathematical
problem is fun and important, in this work, we will focus on the
second part, research and innovation in the hopes of leveraging the
new knowledge in other fields. While we also optimize on a problem,
that problem is not related to a game itself but rather on a meta
level. We try to generate platforming levels, optimizing them to be
similar to existing levels known to the system. Our system will
analyze a dataset of levels and try to generate new levels based on
the seen patterns. It will work without any prior knowledge or feature
engineering done by a human.


%%% Local Variables:
%%% mode: latex
%%% TeX-master: "../../SMWLevelGenerator"
%%% End:

