% Roman numbering for introductory pages
\pagenumbering{roman}

\maketitle

% \newpage

\setcounter{page}{2}
\thispagestyle{empty}
\noindent
\textbf{Declaration of Authorship} \medskip

\noindent
I hereby declare that this master's thesis was authored solely by
myself, Jan~Ebert. All figures and other illustrations were created by
myself unless noted otherwise. All sources and any parts thereof~--
including program source code, data, tables and figures~-- used in
this work have been properly cited and referenced.

\bigskip
\bigskip
\noindent
\makebox[3cm]{\hrulefill} \hspace{0.2cm} \makebox[4cm]{\hrulefill}

\noindent
\makebox[3cm][l]{\textit{Signature}} \hspace{0.2cm} \makebox[4cm][l]{\textit{Place and date}}

% \newpage

\begin{abstract}
  \setcounter{page}{3} We present \emph{SMWLevelGenerator}, a simple
  and extensible preprocessing, training and execution pipeline for
  Super Mario World level generation. We start by generating parts of
  a level via generative adversarial networks, predict metadata via
  image processing-based methods and ``generate'' the rest of the
  level via sequence prediction. We implement a multitude of models, a
  pipeline handling different kinds of training data and easily
  extensible interfaces to handle both more data and more models. The
  training pipeline is optimized for both CPU as well as GPU usage and
  our preprocessing pipeline keeps database sizes to reasonable
  limits.

  Our sequence prediction models learn to generate levels based on the
  most general way possible, meaning they are unplayable and bland.
  While our trained networks do not achieve the desired results, we
  hope to provide an environment for future research to intuitively
  use, improve and extend our collected knowledge.

  A repository containing this text and the source code is available
  at \url{https://github.com/janEbert/SMWLevelGenerator}.
\end{abstract}

% \newpage

\setcounter{page}{4}
\thispagestyle{empty}
\tableofcontents

\newpage

% Restore arabic page numbering
\pagenumbering{arabic}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "../SMWLevelGenerator"
%%% End:

