\subsection{Predicting Metadata}

This section lists implementation details for the final type of model
in this thesis: the metadata predictor. As mentioned in
section~\ref{sec:image-processing}, just like for GANs, there are
convolutional and fully connected models. These models receive the
same inputs as the discriminators described in
section~\ref{sec:first-screen-generation}. Similarly, the MLP-based
model correctly reshapes its inputs. \\
The convolutional metadata predictors were all manually created due to
the mix of convolutional and pooling layers. We use maximal pooling
layers, each followed by a dropout layer, throughout the network but
opt for average pooling as the last pooling operation before the fully
connected output layer.

Our models give as output the constant part of the inputs supplied to
the sequence predictors. Additionally, they output the bit which
indicates whether the level ends at this screen (also an input to the
sequence predictor; explained in
section~\ref{sec:generation-via-prediction}).

The convolutional metadata predictors have the following parameters:
\begin{itemize}
\item starting amount of features in each convolutional layer (these
  are multiplied by powers of two, first increasing, then decreasing.
  Due the models being hand-crafted, there is no clear formula; we
  therefore suggest a curious reader to take a look at the source
  code.)
\item probability of dropout
\item final activation function
\item kernel window size
\end{itemize}
The parameters of the MLP-based metadata prediction models are the
same as those for the MLP-based discriminators in
section~\ref{sec:first-screen-generation}.
As for all models, both the input and output size may be adjusted as well.

We have now not only described our methods in theory but also laid
them out practically, showing which options a casual user has and how
they affect the models internally. Next, we will combine the methods
to get our level generation pipeline.


%%% Local Variables:
%%% mode: latex
%%% TeX-master: "../../SMWLevelGenerator"
%%% End:

